Funding programmes for early career researchers do/not influence their independence

### 2023-08-16 article.Rmd
updated database + added manually searched data for supervisors

---
title: "Funding programmes for early career researchers do/not influence their independence"
author: "David Janků"
date: "`r Sys.Date()`"
output: html_document
bibliography: researcher_independence.bib
csl: apa.csl
---

@@

We used Genderize.io API to estimate the probability of perceived gender based on the first names of researchers. Where available, information about the nationality of researchers was included in the API call to improve accuracy.

@@

For the construction of the control group, we implemented propensity matching using the package MatchIt (citation?) and matching exactly by the discipline (42 sub-categories of OECD classification), and approximately by number of publications in total, number of publications in WoS/Scopus, career age, interdisciplinary proportion, number of grants received prior to Junior Grant, career age when receiving the first grant and gender. All of these variables were calculated at the intervention year (i.e. year when the given researcher from experimental group received the Junior Grant). The distance was set to "mahalanobis". The matching was done in 1:1 ratio and with replacement.

@@

After matching, all standardized mean differences for the covariates were below 0.1 and all standardized mean differences for squares and two-way interactions between covariates were below .15, indicating adequate balance. __This removed XXXX cases from the experimental group, leaving us with YYY matched observations__.

@@

# Run regression model with interaction term
did_model <- plm(researcher_independence ~ independence_timing * treatment + factor(subclass), 
                 data = tar_read(full_indicator), 
                 index = c("subclass", "independence_timing"),
                 model = "within")

# Calculate clustered standard errors
clustered_se <- coeftest(did_model, vcov = function(x) vcovHC(x, type = "HC1", cluster = "group"))
print(clustered_se)

>>>>

---
title: "Funding programmes for early career researchers do/not influence their independence"
author: "David Janků"
date: "`r Sys.Date()`"
output:
  html_document: default
  word_document: default
bibliography: researcher_independence.bib
csl: apa.csl
---

@@

We used Genderize.io API to estimate the probability of perceived gender based on the first names of researchers. Where available, information about the nationality of researchers was included in the API call to improve accuracy. 
We have also manually specified gender for 2 obrsevtation from the treatment group for which the gender was not calculated automatically.

@@

For the construction of the control group, we implemented propensity matching using the package MatchIt (citation?) and matching exactly by the discipline (42 sub-categories of OECD classification) and treatment year, and approximately by number of publications in total, number of publications in WoS/Scopus, career age, interdisciplinary proportion, number of grants received prior to Junior Grant, career age when receiving the first grant and gender. All of these variables were calculated at the intervention year (i.e. year when the given researcher from experimental group received the Junior Grant). The distance was set to "mahalanobis". The matching was done in 1:3 ratio and with replacement. Manual search for supervisors of the matched control group observations was done by ranking the control group observation by distance from the treatment group observation and first searching for the supervisor of the closest control group observation, and if it was not possible to find it, then repeat the same for the second closest, etc... In the end each treatment observation was matched with exactly one control group observation which was closest but at the same time for which it was possible to identify their supervisor.

@@

removed_treatment_unfunded <- (tar_read(ids_complete) %>% distinct(vedidk) %>% nrow())-(tar_read(matching) %>% group_by(subclass) %>% filter(!any(treatment == 2)) %>%
  ungroup() %>% filter(treatment == "1") %>% distinct(vedidk) %>% nrow())

removed_treatment_funded <- (tar_read(ids_complete) %>% distinct(vedidk) %>% nrow())-(tar_read(matching) %>% group_by(subclass) %>% filter(!any(treatment == 0)) %>%
  ungroup() %>% filter(treatment == "1") %>% distinct(vedidk) %>% nrow())

unique_treatment_unfunded <- d %>% group_by(subclass) %>% filter(!any(treatment == 2)) %>% ungroup() %>% filter(treatment == "1") %>% distinct(vedidk) %>% nrow()

unique_treatment_funded <- d %>% group_by(subclass) %>% filter(!any(treatment == 0)) %>%
  ungroup() %>% filter(treatment == "1") %>% distinct(vedidk) %>% nrow()

unique_control_unfunded <- d %>% filter(treatment == 0) %>% distinct(vedidk) %>% nrow()

unique_control_funded <- d %>% filter(treatment == 2) %>% distinct(vedidk) %>% nrow()

replaced_control_unfunded <- unique_treatment_unfunded-unique_control_unfunded

replaced_control_funded <- unique_treatment_funded-unique_control_funded

overlap_control <- unique_control_unfunded+unique_control_funded-(d %>% filter(treatment != 1) %>% distinct(vedidk) %>% nrow())

After matching, all standardized mean differences for the covariates were below 0.1 and all standardized mean differences for squares and two-way interactions between covariates were below .15, indicating adequate balance.

Matching process removed `r removed_treatment_unfunded` treatment observations from matching with unfunded control group and `r removed_treatment_funded` treatment observations from matching with funded control group. This observation was removed because the researchers had no publications prior to the year of application for the grant, which would make it not possible to calculate the independence score for them anyway.

The resulting sample had `r unique_treatment_unfunded` unique treatment group observations paired with `r unique_control_unfunded` unique unfunded control group observations suggesting that `r replaced_control_unfunded` observations from unfunded control group were matched to multiple treatment group observations. Similarly, there were `r unique_treatment_funded` unique treatment group observations paired with `r unique_control_funded` unique funded control group observations, suggesting that `r replaced_control_funded` funded control group observations were matched to multiple treatment group observations.

BTW: there is an overlap between control group observations - `r overlap_control` observations from the unfunded control group also appears in the funded control group

@@

# # Run regression model with interaction term
# did_model <- plm(researcher_independence ~ independence_timing * treatment + factor(subclass), 
#                  data = tar_read(full_indicator), 
#                  index = c("subclass", "independence_timing"),
#                  model = "within")
# 
# # Calculate clustered standard errors
# clustered_se <- coeftest(did_model, vcov = function(x) vcovHC(x, type = "HC1", cluster = "group"))
# print(clustered_se)
